---
title: "VT Atlas of Life"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    social: ["twitter","facebook"]
    theme:
      version: 4
      navbar-bg: "#228B22"
runtime: shiny
---

```{r global, include=FALSE}
library(flexdashboard)
library(shiny)
library(shinydashboard)
library(shinythemes)
library(raster)
library(sf)
library(gdalUtils)
library(DBI)
library(RPostgres)
library(rpostgis)
library(dplyr)
library(dbplyr)
library(VTatlas)
library(mapview)
library(leaflet)
library(leafpop)
library(shinyWidgets)
library(lwgeom)
library(qdap)
library(shinycssloaders)
library(DT)
```

```{r child="pages/00_Input_Data.Rmd"}

```

# Home {data-icon="fa-list"}

```{r HomePageFigure, child="pages/01_Home_Page.Rmd", out.width = '95%', out.height='95%'}

```

# Spatial {data-icon="fa-map" data-navmenu="Overview"}

```{r SpatialOverview, child="pages/02_Spatial_Overview.Rmd"}

```

# State of Biodiversity {data-icon="fa-tree" data-navmenu="Overview"}

```{r StateOfBiodiversity, child="pages/03_State_Of_Biodiversity.Rmd"}

```

# Climate Change {data-icon="fa-globe-americas"}

```{r ClimateChange, child="pages/04_Climate_Change.Rmd"}

```

# Species Richness {data-navmenu="Species Summary"}

```{r SpeciesRichness, child="pages/05_Species_Richness.Rmd"}

```

# Mission species {data-navmenu="Species Summary"}

```{r MissionSpecies, child="pages/06_Mission_Species.Rmd"}

```

# Play with data 

```{r GetDataFromPoint, child="pages/07_Play_With_Data.Rmd"}

```


# Conservation Portal {data-icon="fa-ruler"}

```{r ConservationPortal, child="pages/08_Conservation_Portal.Rmd"}
