---
output: html_document
runtime: shiny
---

## Column {.sidebar}

<br> <br>

<h3>

Welcome to the Vermont Atlas of Life's Dashboard

</h3>


The data displayed on this website were generated by community
scientists that submitted wildlife observations to either <a href="https://www.inaturalist.org/projects/vermont-atlas-of-life" blank = TRUE>iNaturalist</a>,
<a href="https://ebird.org/vt/home" blank = TRUE>eBird</a>, <a href="https://www.e-butterfly.org/newebapp/en/projects/view/22" blank = TRUE>eButterfly</a>, or are from museum specimens collected within
Vermont. <br>

<h4>

What's this dashboard all about?

</h4>

```{r}
renderGauge({
  rate <- thisYrCount/lastYrCount*100
  gauge(rate, min = 0, max = 100, symbol = '%', gaugeSectors(
    success = c(80, 100), warning = c(40, 79), danger = c(0, 39)
  ))
})
```

## Column {data-width="800"}

The map below shows all observations submitted to community science
platforms such as iNaturalist, eBird, eButterfly.

```{r}
#https://data-blog.gbif.org/post/gbif-maps-api-using-r-and-leaflet/

# need to define new projection. Only this projection will work with custom queries. 
epsg4326 <- leafletCRS(crsClass = "L.CRS.EPSG4326", code = "EPSG:4326",
            proj4def = "+proj=longlat +datum=WGS84 +no_defs",
            resolutions = 2^(10:0),
            origin =c(0,0)
            )
# create the gbif-geyser style raster layer 
projection <- '4326' # must use this projection code for custom maps
style <- 'style=gbif-classic' # I think any style will work 

prefix = 'https://api.gbif.org/v2/map/occurrence/density/{z}/{x}/{y}@3x.png?'

style = 'style=purpleYellow.point'

gbiftile = paste0(prefix,style)
ebirdtile = paste0(prefix,
                   'style=glacier.point',
                   "&",'datasetKey=4fa7b334-ce0d-4e88-aaae-2e0c138d049e') # eBird key

iNattile = paste0(prefix,
                   'style=fire.point',
                   "&",'datasetKey=50c9509d-22c7-4a22-a47d-8c48425ef4a7') # iNat key

ebutttile <- paste0(prefix,
                    'style=purpleYellow.point',
                    "&",'datasetKey=cf3bdc30-370c-48d3-8fff-b587a39d72d6') #eButt key

leaflet(VTshape) %>%
  addProviderTiles("CartoDB.Positron",
                   group = "baselayer") %>%
  addPolygons(color = "black",
              fillColor = "transparent",
              weight = 1, 
              smoothFactor = 0.5,
              highlightOptions = highlightOptions(color = "white", 
                                                  weight = 2,
                                                  bringToFront = TRUE) 
  ) %>%
  setView(lng=-73,lat=44,zoom=08) %>% 
addPolygons(color = "black",
            fillColor = "transparent",
            weight = 1, 
            smoothFactor = 0.5,
            highlightOptions = highlightOptions(color = "white", 
                                                weight = 2,
                                                bringToFront = TRUE)
           ) %>%
addTiles(urlTemplate=gbiftile,
         group = "GBIF") %>% 
addTiles(urlTemplate=ebirdtile,
         group = "eBird") %>%
addTiles(urlTemplate=iNattile,
         group = "iNaturalist") %>%
addTiles(urlTemplate=ebutttile,
         group = "eButterfly") %>%
  
hideGroup(c("GBIF","eButterfly")) %>%

addLayersControl(
    baseGroups = c("baselayer"),
    overlayGroups = c("GBIF","eBird","iNaturalist","eButterfly"),
    options = layersControlOptions(collapsed = FALSE)
  )  
```

