---
output: html_document
runtime: shiny
---

## Import shapefiles {.sidebar}

<br> Add a shapefile for a region of interest within Vermont. Files with
the following extensions are needed .shp,.dbf, it's recommended to have
.prj as well. <br>

```{r ConservationShaplefileInputs}
fileInput("shp", 
          "Shapefile Import", 
          multiple = TRUE, 
          accept = c('.shp', '.dbf','.sbn', '.sbx', '.shx', '.prj'))

sfuploaded <- eventReactive(input$shp,{
    req(input$shp)
    # create a temp directory to store the files
    tmpdir <- tempdir()
    # delete that folder after exiting 
    on.exit(unlink(tmpdir))
    
    #filename <- substr(input$shp$name[1],1,nchar(input$shp$name[1])-4)
    file.copy(input$shp$datapath,file.path(tmpdir,input$shp$name) )
    upshape <- sf::read_sf(dsn = paste0(tmpdir,"/",input$shp$name[grep(pattern = '.shp', input$shp$name)]),
                quiet = TRUE)
    upshape <- sf::st_transform(upshape, 4326)
  })

selectInput("wants",
          h3("Select what to return"),
          choices = c("Observed species",
                    "Observed richness",
                    "Number of observations"),
          selected = "Observed species")

```

## Show map {data-width="400"}

Ideally, on this page users will be able to upload or import spatial
layers or draw on the map regions of interest to 1) get an 'observed
species list', a 'species that are likely' and future projections.

```{r ConservationShapefilePlot}
renderLeaflet({ 
  M1 <- mapview(VTshape,
               layer.name = "VT boundry",
               col.regions = "transparent",
               alpha.regions = 0.25)
  M2 <- mapview(sfuploaded(),
                layer.name = input$shp$name[grep(pattern = '.shp', input$shp$name)],
                col.regions = "blue"
                )
  (M1+M2)@map
})
```

<b>Note:</b> These queries take a long time and currently, there visual cues
that the process is working. This is a work in progress, when finished a table 
is returned below this message.

```{r ConservationResultTable}
resultTable <- eventReactive(input$wants,{
  
# simplify spatial layer
sfuploaded_shape <- sf::st_union(sfuploaded())

GBIF_Db <- DBI::dbConnect(RPostgres::Postgres(),
                         dbname = 'gis',
                         host = 'dashboard-db.vtatlasoflife.org',
                         port = 6543, # or any other port specified by your DBA 
                         user = "readonly",
                         password = "webaccess")

if(input$wants == "Observed species"){
 
dbGetQuery(GBIF_Db,
           paste0(
             'SELECT DISTINCT \"species\"
             FROM occurrence
             WHERE st_intersects(occurrence.geometry,\'SRID=4326;',
             sf::st_as_text(sfuploaded_shape),'\');'))
}
if(input$wants == "Observed richness"){
 
dbGetQuery(GBIF_Db,
           paste0(
             'SELECT \"order\", COUNT(DISTINCT \"species\")
             FROM occurrence
             WHERE st_intersects(occurrence.geometry,\'SRID=4326;',
             sf::st_as_text(sfuploaded_shape),'\')
             GROUP BY \"order\";'))
}
if(input$wants == "Number of observations"){
 
dbGetQuery(GBIF_Db,
           paste0(
             'SELECT COUNT(DISTINCT \"gbifID\")
             FROM occurrence
             WHERE st_intersects(occurrence.geometry,\'SRID=4326;',
             sf::st_as_text(sfuploaded_shape),'\');'))
}  
})

renderDataTable(
                ifelse(input$wants=="Number of observations",
                       as.numeric(resultTable()),
                       DT::datatable(resultTable(),
                options=list(
                  lengthMenu= list(c(5,25,100,-1), 
                                   c('5','25','100','All')),
                  pageLength = 25)
                )))
```